Concordance between observations and simulations in the evolution of the mass relation between supermassive black holes and their host galaxies

### 2022-05-24 projects/2020_HSC_to_numerical_simulation/manuscript/manuscript.tex
For uploading final materials

\newcommand{\vdag}{(v)^\dagger}
\newcommand\aastex{AAS\TeX}
\newcommand\latex{La\TeX}
%\newcommand{\red}[1]{{#1}}
\newcommand{\red}[1]{\textcolor{red}{#1}}
\newcommand{\blue}[1]{\textcolor{blue}{\bf #1}}

@@

\section{Introduction} \label{sec:intro}
The close correlations between the mass of supermassive black holes (\mbh) and the properties of their host galaxies (e.g., stellar mass, \smass) indicate a physical coupling during their joint evolution~\citep{Mag++98, F+M00, M+H03, H+R04, Gul++09}. To understand the nature of this connection, considerable efforts have been focused on measuring such correlations using broad-line active galactic nucleus (AGN) over a range of redshifts with the intention to determine how and when the correlation emerges and evolves over cosmic time. While, an {\it observed} evolution has been found since redshift $z\sim2$ in galaxies with stellar mass \smass\ $\gtrsim 10^{10}M_{\odot}$ \citep[e.g.,][]{Tre++04, Peng2006a,Tre++07,Woo++08,Bennert11,Park15} in which the growth of BHs predates that of the host, other studies\blue{~\citep[e.g.,][]{Schramm2013, Sun2015, 2021ApJ...906..103L}} predict that BHs grow commensurately with galaxies.
However, to understand the significance {\it intrinsic} evolution, it is necessary to take into account systematic uncertainties and the selection effects~\citep{Tre++07, Lauer2007, Schulze2014, Park15,Jahnke2009,  Ding2020,Li2021b}.

Various theoretical models have been proposed to explain the origin of the scaling relations. For example, AGN feedback is considered as one of the possible viable mechanisms. During this process, a fraction of the AGN energy is injected into its surrounding gas, which can then regulate the mass growth of the BH and its host. In this scenario, star formation is inhibited by the heating and unbinding of a significant amount of gas. Alternatively, the mass relations can be explained through an indirect connection in which AGN accretion and star formation are fed through a common gas supply~\citep{Cen2015, Menci2016, angles_black_2017}. Actually, even without any physical mechanisms, statistical convergence from galaxy assembly alone (i.e., dry mergers) could instill the observed correlations~\citep{Peng2007, Jahnke2011, Hirschmann2010}. However, as expected from the central limit theorem, a higher dispersion would appear in the scaling relations at high-$z$ compared to what is observed today~\blue{\citep[e.g.,][]{Ginat2016, 2020MNRAS.498.5652K}}.

Numerical simulations provide an opportunity to further understand the connection between BHs and their host galaxies. For example, a comparison of scaling relations has been made using state-of-the-art cosmological hydrodynamical simulation of structure formation ({\tt MassiveBlackII}) and observational measurements at  $0.3<z<1$ \citep[e.g., ][]{DeG++15}, which show a positive evolution (i.e., the \blue{mass growth of the BH predates} that of its host). Further efforts are using large-volume simulations to investigate the scaling relations and find good agreement with the local relation with redshift evolution, 
including the Magneticum Pathfinder smooth-particle hydrodynamics (SPH) Simulations~\citep{Steinborn2015}, the Evolution and Assembly of Galaxies and their Environments~(EAGLE) suite of SPH simulations~\citep{Schaye2015, Crain2015, McAlpine2016}, Illustris moving-mesh simulation~\citep{Genel2014, 2014MNRAS.444.1518V, Sijacki2015, Nelson2015, Li2019}, the Horizon-AGN simulation~\citep{2014MNRAS.444.1453D, 2016MNRAS.463.3948D, 2016MNRAS.460.2979V}, and the SIMBA simulation~\citep{Thomas2019, Dave2019}.
In particular, the \mbh-\smass\ relation using BH populations using six large-scale cosmological simulations (i.e, Illustris, TNG100, TNG300, Horizon-AGN, EAGLE, and SIMBA) has been compared with observations in the local universe in a recent study~\citep{Habouzit2021}. However, these comparison works are limited by the observation data in terms of the sample size ($<$100) and redshift range (i.e., limited to the local universe).

@@

\subsection{SDSS/HSC sample}\label{sec:hsc}
A sample of $\sim$5000 type-1 SDSS quasars from the DR14 catalog~\citep{Paris2018} at $0.2<z<1$ has been imaged by the high-resolution Subaru Strategic Program (SSP) wide area survey~\citep{Aihara2019} using Hyper Suprime-Cam~\citep{Miyazaki2018}. With accurate PSF models in five optical bands {\it grizy}, two-dimensional quasar-host decompositions have been performed \citep[][]{Li2021a} to obtain the flux and color of each quasar's host galaxy. The \blue{state-of-the-art} image modeling software {\tt lenstronomy} \citep{Birrer2015, Birrer2018, Birrer2021} is adopted to perform the modeling task. This approach is first developed by~\citet{Ding2020} and used to decompose the near-infrared emission of the HST sample (see next section). Having measured the host light in each band, the stellar mass of host galaxy is derived using spectral energy distribution (SED) fitting with CIGALE~\citep{Boquien2019}. Simulation tests are also performed to verify the fidelity of the \smass\ measurements. The statistical measurement error on \smass\ is at the $\sim$0.2~dex level. The values of \mbh\ are determined by~\citet{Rakshit2020} which are estimated based on the \hbeta-based measurements using the virial method~\citep{Peterson2004, Vestergaard2006}. The typical error of \mbh\ are estimated to be 0.4 dex. The mass ranges for the entire sample are log(\smass) $\in$ [9.0, 11.5] $M_{\odot}$, and log(\mbh) $\in$  [6.5, 10.0] $M_{\odot}$).
 We refer the reader to~\citet{Li2021a} in the Section~4.2 for more details.

@@

To model supermassive BH, the initial seed with mass $5\times 10^{5}~M_{\odot}/h$ \blue{is} inserted into halos of mass $\gtrsim 5\times 10^{10}~M_{\odot}/h$. Once seeded, BH growth via gas accretion is assigned at a rate of $\dot{M}_{BH}={4\pi G^2 M_{BH}^2 \rho}/{(c_s^2+v_{BH}^2)^{3/2}}$ where $\rho$ and $c_s$ are the density and sound speed of the interstellar medium (ISM) gas at cold phase; $v_{BH}$ is the relative velocity between the BH and its surrounding gas. Note that unlike several previous works, the accretion rate in MBII adopt the prescription in~\citet{Pelupessy2007} which does not include any artificial boost factor. The accreted gas is released as radiation at a radiative efficiency of 10\%. A fraction of 5\% of the radiated energy thermally couples to the surrounding gas as BH (or AGN) feedback~\citep{2005Natur.433..604D}. A mildly super-Eddington (two times Eddington rate) is allowed. Due to resolution limitations, BH dynamics cannot be self-consistently modeled in the simulations. Two BHs are considered to be merged when their separation distance is below the simulation spatial resolution (i.e., the SPH smoothing length) and their relative speeds are lower than the local sound speed of the medium.

@@

In Figure~\ref{fig:comparsion}, we present the mass scaling relation \mbh--\smass\ for both the observations and simulations for direct comparison. For the simulated data, both the initial sample and that with observational effects applied (i.e., noise correction, and selection) are shown. Note that offsets can occur for the simulated \blue{samples} in three cases: an inherent offset, an offset due to selection, and offsets from added noise and selection. For the first, it has been recognized~\citep[e.g.,][Figure 2]{Habouzit2021}, and seen in our Figure~\ref{fig:comparsion}, that the  offset values vary over a range of stellar mass, and thus the mean and standard derivation do not represent the entire mass distribution.
In this work, we focus on the latter two offset distributions, i.e., considering selection effects with and without noise correction. To aid in visualization of the differences among the various simulations, compared to the observed sample, we show the distribution of offsets (in terms of the $\Delta{\rm log}$\mbh) as histograms in Figure~\ref{fig:offsets}.  Each panel presents a different redshift range. In addition, the values of the central offset and scatter are presented 
in Table~\ref{tab:sum}, both before and after consideration of the noise. Comparisons with the observations are presented in the remainder of this section.

\begin{figure*}
\centering
\includegraphics[trim = 40mm 60mm 30mm 90mm, clip,height=1.1\textwidth]{MM_sum.png}
\caption{\label{fig:comparsion} 
Black hole mass versus stellar mass for both the observational (small orange circles) and simulated (small colored circles) samples. Each row pertains to a particular simulation as labelled. The panels, from left to right, show different redshift bins. The black line in each panel indicates the local relation adopted by~\citet{Ding2020}. The background cloud (in green and yellow with contours) shows the intrinsic simulation number density before injecting random noise and applying selection effects. The TNG100 and TNG300 appear to present similar results (see discussion section). \blue{Note that different sample adopt either a Chabrier or a Salpeter IMF for calculating the stellar mass, thus the \smass\ value for the observation and the local relation can be slightly different in each vertical plane.}
}
\end{figure*}

@@

\section{Discussion} \label{sec:dis}
In this study, there are a few issues that may bias the results thus need to be considered. First, the mass offsets are compared to the \blue{observed} relation derived in the local Universe (D20). In fact, the different simulations could have different mean relations at $z=0$~\citep[e.g.,][]{Habouzit2021}. As a result, the interpretation of the BH mass offsets of the simulations with redshift, anchored to the D20 local scaling relation, can be different if using the local relation of each individual simulation. In addition, the stellar mass of the Horizon-AGN sample is the total mass while that for the other hydrodynamic simulations are determined within a 3D 30 kpc spherical aperture\footnote{\blue{In fact, we also checked to use total stellar mass for Illustris, TNG100, and TNG300 for the comparison, the results appear to be consistent with the ones when using the 30 kpc stellar mass.}}. Therefore, in Figure~\ref{fig:offsets_vz_shiftz03}, we manually \blue{recalibrate the offset of all simulations so that their mean offset value  at $z=0.3$ are at the same level with the observations}. This enables the evolutionary trend in the offsets of each sample to be clearer. In general, the updated results show that the evolution of the simulations are consistent with our interpretations in Section~\ref{subsec:offset}. This result can be expected since most of the simulations have demonstrated a good match to the observations at redshift $z=0.3$.

Regarding the survey volumes, we collected all available samples in the simulations to perform these comparisons. The final sample sizes, used to compare with observation, are not set by design. Despite that the observed samples are not volume-limited and the simulated samples are not volume-matched, fortunately, the volume of all the simulations appear to be sufficient, i.e., all simulations have a similar amount (or more) of data points to compare with the observations (see Figure~\ref{fig:comparsion}). Also, similar results are found with the smaller and larger volumes simulations, e.g., TNG100 and TNG300, which reflect the fact that the simulation volume reaches a sufficient size to be effective for \blue{the comparisons in the observational plane. Even though, the intrinsic scatters of TNG100 and TNG300 (before the noise/selection) can be still different.}\footnote{\blue{In particular, our study is made at \smass$>10~M_{\odot}$, while intrinsically, the subgrid models of the TNG simulation can change a lot at low redshift, especially for \smass$<10.5~M_{\odot}$ (see green controus in Figure 3 for $z<0.7$ sample). }}

@@

\begin{deluxetable*}{lcccc}
\tablecaption{Summary of the p-value using KS test \label{tab:pvalue}}
\tablewidth{0pt}
\tablehead{
\colhead{Simulation} &  \multicolumn{3}{c}{HSC comparison} & \colhead{HST comparison} \\
\cline{2-4} 
\colhead{} &  \colhead{$z\sim0.3$} & \colhead{$z\sim0.5$} & \colhead{$z\sim0.7$} & \colhead{$z\sim1.5$}
}
\startdata
SAM &  1.99e-09 & $<$1e-10  & $<$1e-10  & 7.46e-03  \\
MBII & 7.51e-01 & \multicolumn{2}{c}{$<$3.93e-01  [$z=0.6$]}  & 2.22e-05  \\
Illustris & 5.01e-01 & 8.38e-01  & 5.98e-01  & 3.90e-05  \\
TNG100 & 1.15e-02 & 4.06e-01  & 4.84e-01  & 4.71e-01  \\
TNG300 & 2.36e-02 & 5.46e-01  & 3.10e-01  & 2.12e-01  \\
Horizon-AGN & 2.01e-01 & 4.83e-01  & 1.40e-01  & 1.95e-01  \\
\enddata
\tablecomments{These p-values are obtained by the KS test between the simulation and the observation based on one realization.}
\end{deluxetable*}

\bibliography{reference}{}
%\input{manuscript.bbl}
\bibliographystyle{aasjournal}

>>>>

\newcommand{\vdag}{(v)^\dagger}
\newcommand\aastex{AAS\TeX}
\newcommand\latex{La\TeX}
%\newcommand{\red}[1]{{#1}}
%\newcommand{\red}[1]{\textcolor{red}{#1}}
%\newcommand{\blue}[1]{\textcolor{black}{\bf #1}}

@@

\section{Introduction} \label{sec:intro}
The close correlations between the mass of supermassive black holes (\mbh) and the properties of their host galaxies (e.g., stellar mass, \smass) indicate a physical coupling during their joint evolution~\citep{Mag++98, F+M00, M+H03, H+R04, Gul++09}. To understand the nature of this connection, considerable efforts have been focused on measuring such correlations using broad-line active galactic nucleus (AGN) over a range of redshifts with the intention to determine how and when the correlation emerges and evolves over cosmic time. While, an {\it observed} evolution has been found since redshift $z\sim2$ in galaxies with stellar mass \smass\ $\gtrsim 10^{10}M_{\odot}$ \citep[e.g.,][]{Tre++04, Peng2006a,Tre++07,Woo++08,Bennert11,Park15} in which the growth of BHs predates that of the host, other studies~\citep[e.g.,][]{Schramm2013, Sun2015, 2021ApJ...906..103L} predict that BHs grow commensurately with galaxies.
However, to understand the significance {\it intrinsic} evolution, it is necessary to take into account systematic uncertainties and the selection effects~\citep{Tre++07, Lauer2007, Schulze2014, Park15,Jahnke2009,  Ding2020,Li2021b}.

Various theoretical models have been proposed to explain the origin of the scaling relations. For example, AGN feedback is considered as one of the possible viable mechanisms. During this process, a fraction of the AGN energy is injected into its surrounding gas, which can then regulate the mass growth of the BH and its host. In this scenario, star formation is inhibited by the heating and unbinding of a significant amount of gas. Alternatively, the mass relations can be explained through an indirect connection in which AGN accretion and star formation are fed through a common gas supply~\citep{Cen2015, Menci2016, angles_black_2017}. Actually, even without any physical mechanisms, statistical convergence from galaxy assembly alone (i.e., dry mergers) could instill the observed correlations~\citep{Peng2007, Jahnke2011, Hirschmann2010}. However, as expected from the central limit theorem, a higher dispersion would appear in the scaling relations at high-$z$ compared to what is observed today~\citep[e.g.,][]{Ginat2016, 2020MNRAS.498.5652K}.

Numerical simulations provide an opportunity to further understand the connection between BHs and their host galaxies. For example, a comparison of scaling relations has been made using state-of-the-art cosmological hydrodynamical simulation of structure formation ({\tt MassiveBlackII}) and observational measurements at  $0.3<z<1$ \citep[e.g., ][]{DeG++15}, which show a positive evolution (i.e., the mass growth of the BH predates that of its host). Further efforts are using large-volume simulations to investigate the scaling relations and find good agreement with the local relation with redshift evolution, 
including the Magneticum Pathfinder smooth-particle hydrodynamics (SPH) Simulations~\citep{Steinborn2015}, the Evolution and Assembly of Galaxies and their Environments~(EAGLE) suite of SPH simulations~\citep{Schaye2015, Crain2015, McAlpine2016}, Illustris moving-mesh simulation~\citep{Genel2014, 2014MNRAS.444.1518V, Sijacki2015, Nelson2015, Li2019}, the Horizon-AGN simulation~\citep{2014MNRAS.444.1453D, 2016MNRAS.463.3948D, 2016MNRAS.460.2979V}, and the SIMBA simulation~\citep{Thomas2019, Dave2019}.
In particular, the \mbh-\smass\ relation using BH populations using six large-scale cosmological simulations (i.e, Illustris, TNG100, TNG300, Horizon-AGN, EAGLE, and SIMBA) has been compared with observations in the local universe in a recent study~\citep{Habouzit2021}. However, these comparison works are limited by the observation data in terms of the sample size ($<$100) and redshift range (i.e., limited to the local universe).

@@

\subsection{SDSS/HSC sample}\label{sec:hsc}
A sample of $\sim$5000 type-1 SDSS quasars from the DR14 catalog~\citep{Paris2018} at $0.2<z<1$ has been imaged by the high-resolution Subaru Strategic Program (SSP) wide area survey~\citep{Aihara2019} using Hyper Suprime-Cam~\citep{Miyazaki2018}. With accurate PSF models in five optical bands {\it grizy}, two-dimensional quasar-host decompositions have been performed \citep[][]{Li2021a} to obtain the flux and color of each quasar's host galaxy. The state-of-the-art image modeling software {\tt lenstronomy} \citep{Birrer2015, Birrer2018, Birrer2021} is adopted to perform the modeling task. This approach is first developed by~\citet{Ding2020} and used to decompose the near-infrared emission of the HST sample (see next section). Having measured the host light in each band, the stellar mass of host galaxy is derived using spectral energy distribution (SED) fitting with CIGALE~\citep{Boquien2019}. Simulation tests are also performed to verify the fidelity of the \smass\ measurements. The statistical measurement error on \smass\ is at the $\sim$0.2~dex level. The values of \mbh\ are determined by~\citet{Rakshit2020} which are estimated based on the \hbeta-based measurements using the virial method~\citep{Peterson2004, Vestergaard2006}. The typical error of \mbh\ are estimated to be 0.4 dex. The mass ranges for the entire sample are log(\smass) $\in$ [9.0, 11.5] $M_{\odot}$, and log(\mbh) $\in$  [6.5, 10.0] $M_{\odot}$).
 We refer the reader to~\citet{Li2021a} in the Section~4.2 for more details.

@@

To model supermassive BH, the initial seed with mass $5\times 10^{5}~M_{\odot}/h$ is inserted into halos of mass $\gtrsim 5\times 10^{10}~M_{\odot}/h$. Once seeded, BH growth via gas accretion is assigned at a rate of $\dot{M}_{BH}={4\pi G^2 M_{BH}^2 \rho}/{(c_s^2+v_{BH}^2)^{3/2}}$ where $\rho$ and $c_s$ are the density and sound speed of the interstellar medium (ISM) gas at cold phase; $v_{BH}$ is the relative velocity between the BH and its surrounding gas. Note that unlike several previous works, the accretion rate in MBII adopt the prescription in~\citet{Pelupessy2007} which does not include any artificial boost factor. The accreted gas is released as radiation at a radiative efficiency of 10\%. A fraction of 5\% of the radiated energy thermally couples to the surrounding gas as BH (or AGN) feedback~\citep{2005Natur.433..604D}. A mildly super-Eddington (two times Eddington rate) is allowed. Due to resolution limitations, BH dynamics cannot be self-consistently modeled in the simulations. Two BHs are considered to be merged when their separation distance is below the simulation spatial resolution (i.e., the SPH smoothing length) and their relative speeds are lower than the local sound speed of the medium.

@@

In Figure~\ref{fig:comparsion}, we present the mass scaling relation \mbh--\smass\ for both the observations and simulations for direct comparison. For the simulated data, both the initial sample and that with observational effects applied (i.e., noise correction, and selection) are shown. Note that offsets can occur for the simulated samples in three cases: an inherent offset, an offset due to selection, and offsets from added noise and selection. For the first, it has been recognized~\citep[e.g.,][Figure 2]{Habouzit2021}, and seen in our Figure~\ref{fig:comparsion}, that the  offset values vary over a range of stellar mass, and thus the mean and standard derivation do not represent the entire mass distribution.
In this work, we focus on the latter two offset distributions, i.e., considering selection effects with and without noise correction. To aid in visualization of the differences among the various simulations, compared to the observed sample, we show the distribution of offsets (in terms of the $\Delta{\rm log}$\mbh) as histograms in Figure~\ref{fig:offsets}.  Each panel presents a different redshift range. In addition, the values of the central offset and scatter are presented 
in Table~\ref{tab:sum}, both before and after consideration of the noise. Comparisons with the observations are presented in the remainder of this section.

\begin{figure*}
\centering
\includegraphics[trim = 40mm 60mm 30mm 90mm, clip,height=1.1\textwidth]{MM_sum.png}
\caption{\label{fig:comparsion} 
Black hole mass versus stellar mass for both the observational (small orange circles) and simulated (small colored circles) samples. Each row pertains to a particular simulation as labelled. The panels, from left to right, show different redshift bins. The black line in each panel indicates the local relation adopted by~\citet{Ding2020}. The background cloud (in green and yellow with contours) shows the intrinsic simulation number density before injecting random noise and applying selection effects. The TNG100 and TNG300 appear to present similar results (see discussion section). We note that different samples adopt either a Chabrier or a Salpeter IMF for calculating the stellar mass thus the \smass\ values for the observations and the local relation are shifted appropriately.
}
\end{figure*}

@@

\section{Discussion} \label{sec:dis}
In this study, there are a few issues that may bias the results thus need to be considered. First, the mass offsets are compared to the observed relation derived in the local Universe (D20). In fact, the different simulations could have different mean relations at $z=0$~\citep[e.g.,][]{Habouzit2021}. As a result, the interpretation of the BH mass offsets of the simulations with redshift, anchored to the D20 local scaling relation, can be different if using the local relation of each individual simulation. In addition, the stellar mass of the Horizon-AGN sample is the total mass while that for the other hydrodynamic simulations are determined within a 3D 30 kpc spherical aperture\footnote{We checked the use of total stellar mass for Illustris, TNG100, and TNG300 in these comparisons; the results are consistent with those based on a 30 kpc region.}. Therefore, in Figure~\ref{fig:offsets_vz_shiftz03}, we manually recalibrate the offsets of all simulations so that their mean value at $z=0.3$ is fixed to match that of the observations. This enables the evolutionary trend in the offsets of each sample to be clearer. In general, the updated results show that the evolution of the simulations are consistent with our interpretations in Section~\ref{subsec:offset}. This result can be expected since most of the simulations have demonstrated a good match to the observations at redshift $z=0.3$.

Regarding the survey volumes, we collected all available samples in the simulations to perform these comparisons. The final sample sizes, used to compare with observation, are not set by design. Despite that the observed samples are not volume-limited and the simulated samples are not volume-matched, fortunately, the volume of all the simulations appear to be sufficient, i.e., all simulations have a similar amount (or more) of data points to compare with the observations (see Figure~\ref{fig:comparsion}). Also, similar results are found with the smaller and larger volumes simulations, e.g., TNG100 and TNG300, which reflect the fact that the simulation volume reaches a sufficient size to be effective for the comparisons in the observational plane. Even though, the intrinsic scatter of TNG100 and TNG300 (before the noise/selection) can differ.\footnote{We note that 
%Our study is primarily at \smass$>10~M_{\odot}$, while 
the subgrid models of the TNG simulation can change the intrinsic distributions at low redshift, especially for log~\smass$<10.5~M_{\odot}$ (see green contours in Figure 3 for $z<0.7$ sample). }

@@

\begin{deluxetable*}{lcccc}
\tablecaption{Summary of the p-value using KS test \label{tab:pvalue}}
\tablewidth{0pt}
\tablehead{
\colhead{Simulation} &  \multicolumn{3}{c}{HSC comparison} & \colhead{HST comparison} \\
\cline{2-4} 
\colhead{} &  \colhead{$z\sim0.3$} & \colhead{$z\sim0.5$} & \colhead{$z\sim0.7$} & \colhead{$z\sim1.5$}
}
\startdata
SAM &  1.99e-09 & $<$1e-10  & $<$1e-10  & 7.46e-03  \\
MBII & 7.51e-01 & \multicolumn{2}{c}{3.93e-01  [$z=0.6$]}  & 2.22e-05  \\
Illustris & 5.01e-01 & 8.38e-01  & 5.98e-01  & 3.90e-05  \\
TNG100 & 1.15e-02 & 4.06e-01  & 4.84e-01  & 4.71e-01  \\
TNG300 & 2.36e-02 & 5.46e-01  & 3.10e-01  & 2.12e-01  \\
Horizon-AGN & 2.01e-01 & 4.83e-01  & 1.40e-01  & 1.95e-01  \\
\enddata
\tablecomments{These p-values are obtained by the KS test between the simulation and the observation based on one realization.}
\end{deluxetable*}

%\bibliography{reference}{}
\input{manuscript.bbl}
\bibliographystyle{aasjournal}
